Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down

### 2022-08-26 paper/ms.tex
Move figures around

\section{Introduction}
\label{sec:intro}
Approximately two-thirds of spiral disks host an elongated bar-like feature at
their center \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our own
Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of bars
is not difficult to explain, since stellar disks simulated in isolation almost
always form bar-like structures \citep{1971ApJ...168..343H}. Several studies
have shown that a spherical potential, such as a stellar bulge or dark matter
halo, acts to stabilize a stellar disk against the bar instability
\citep{1973ApJ...186..467O, 1976AJ.....81...30H}.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{fig/mockHST.pdf}
    \caption{Synthetic Hubble Space Telescope images of our simulations with and
    without the interstellar medium. The upper panels show an \Nbody{} only
    simulation while the lower panels show a simulation which includes the
    SMUGGLE model for the interstellar medium. Each panel is $20\,\textrm{kpc}$
    to a side. Columns show different points in time, separated by
    $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
    length and strength. In the SMUGGLE run, the bar remains at approximately
    the same length and strength over the course of the simulation. The \Nbody{}
    model is identical to the GALAKOS model, discussed in the text. Details on
    the production of these synthetic images is given in Appendix~\ref{app:hst}.}\label{fig:overview}
\end{figure*}

@@

\section{Methods}
\label{sec:methods}

\subsection{Initial Conditions}
The initial setup of the galactic disk used in this work follows closely the
GALAKOS model \citep{2020ApJ...890..117D}, which uses a modified version of the
\texttt{MakeNewDisk} code \citep{2005MNRAS.361..776S}. The GALAKOS model has
three components - a radially exponential and vertically isothermal stellar
disk, and a stellar bulge and dark matter halo following a Hernquist profile
\citep{1990ApJ...356..359H}. All \Nbody{} runs in this work used the same setup
parameters as the GALAKOS disk, more details of which can be found in the
original paper.

@@

\section{Results}
\label{sec:results}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{fig/ps_torque.pdf}
    \caption{The \textit{upper panel} shows the evolution
    of the pattern speed. As expected, the bar in the \Nbody{} run slows down due
    to interactions between the bar and the dark matter halo. However, the bar in
    the SMUGGLE run does not slow down and instead remains at a constant pattern
    speed. The \textit{lower panel} shows the torque on the bar
    by different components. The solid lines show the torque exerted by the halo
    in both the \Nbody{} and SMUGGLE cases. The dashed line shows the torque
    exerted by the gas phase in the SMUGGLE run (there is no gas in the \Nbody{}
    run). Details on how these properties are calculated is given in the Methods
    section.}\label{fig:prop}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/Rb_A2.pdf}
    \caption{The evolution in bar length and strength. The \textit{upper panel}
    shows the evolution of the bar length. In the \Nbody{} case, the bar
    lengthens. This occurs because as the pattern speed drops, bar-like orbits
    at larger radii are possible. Stars are captured on these orbits,
    lengthening the bar. This process does not occur in the SMUGGLE cases since
    the bar pattern speed is not decreasing, and therefore the bar length
    remains constant. The bar strength, shown in the \textit{lower panel} is
    measured as the maximum of the second Fourier component divided by the
    zeroth Fourier component. We see that in the \Nbody{} case (blue) the bar
    strength increases with time, consistent with previous results showing that
    the bar strength increases as bars slow down. In the SMUGGLE case (orange),
    we see that the bar strength remains roughly constant, possibly slightly
    decreasing with time. This is also consistent with the expected relation
    between pattern speed and strength since the bar in this case is not slowing
    down.}
    \label{fig:strength}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/ps_late_start.pdf}
    \caption{Pattern speed evolution with a lower initial pattern speed. We
    tested the evolution of our system when gas is added to the \Nbody{} run at
    a later time, but with all other simulation parameters kept the same. The
    setup is therefore identical to our previous runs just with a bar that is
    larger, stronger, and with a lower pattern speed. We find that the pattern
    speed evolution is very similar to our fiducial case, except that the bar
    retains its original pattern speed. This indicates a mechanism which keeps
    the bar at its formation pattern speed, and that there is not a particular
    pattern speed which the system tends to.}
    \label{fig:snap700}
\end{figure}

@@

\begin{figure}[t!]
    \centering
    \includegraphics[width=9cm]{fig/ps_fgas.pdf}
    \caption{Pattern speed evolution with varying gas fractions. We explored the
    impact of lowering the initial gas surface density of our fiducial disk on
    the evolution of the pattern speed. The surface densities we tested of $20$,
    $15$, $10$, and $5\,\Msun/\textrm{kpc}^2$ correspond to initial gas
    fractions of $16\%$, $10\%$, $7\%$, and $4\%$. We find that initial surface
    densities $20$, $15$, and $10\,\Msun/\textrm{kpc}^2$ result in bars which
    remain at a constant pattern speed, while an initial surface density of
    $5\,\Msun/\textrm{kpc}^2$ results in a bar which slows down in roughly the
    same manner as the \Nbody{} case.}
    \label{fig:fgas}
\end{figure}

@@

\begin{figure}[h!]%
    \centering
    \includegraphics[width=9cm]{fig/ps_GFM.pdf}
    \caption{Pattern speed evolution of a smooth ISM model. This evolution is
    shown for the fiducial disk in the \Nbody{} (blue), SMUGGLE (orange), and
    smooth ISM (red) cases. The smooth ISM model is an older model for the ISM
    which treats its multiphase nature in a subgrid fashion
    \citep{2003MNRAS.339..289S}. This fundamentally differs from the SMUGGLE
    model, which explicitly resolves the hot and cold phases of the ISM
    \citep{2019MNRAS.489.4233M}. The pattern speed in the smooth ISM case is
    broadly similar to the evolution in the SMUGGLE case. This shows that the
    stability of the pattern speed is not simply a result of our assumed model
    for the ISM.}
\label{fig:GFM}
\end{figure}

@@

First, we note that \citet{2020MNRAS.491.3655G} found that the rotation
parameter \Rot{} positively correlates with gas fraction, such that galaxies
with higher gas fractions are rotating more slowly. However, it is not obvious
this is in tension with our result since the gas fraction of galaxies correlates
with other galactic properties \citep{2009ARAA..47..159B}. Furthermore, the
measurement of pattern speeds is a delicate process still prone to large errors.

@@

>>>>

\section{Introduction}
\label{sec:intro}
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{fig/mockHST.pdf}
    \caption{Synthetic Hubble Space Telescope images of our simulations with and
    without the interstellar medium. The upper panels show an \Nbody{} only
    simulation while the lower panels show a simulation which includes the
    SMUGGLE model for the interstellar medium. Each panel is $20\,\textrm{kpc}$
    to a side. Columns show different points in time, separated by
    $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
    length and strength. In the SMUGGLE run, the bar remains at approximately
    the same length and strength over the course of the simulation. The \Nbody{}
    model is identical to the GALAKOS model, discussed in the text. Details on
    the production of these synthetic images is given in Appendix~\ref{app:hst}.}\label{fig:overview}
\end{figure*}

Approximately two-thirds of spiral disks host an elongated bar-like feature at
their center \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our own
Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of bars
is not difficult to explain, since stellar disks simulated in isolation almost
always form bar-like structures \citep{1971ApJ...168..343H}. Several studies
have shown that a spherical potential, such as a stellar bulge or dark matter
halo, acts to stabilize a stellar disk against the bar instability
\citep{1973ApJ...186..467O, 1976AJ.....81...30H}.

@@

@@

\section{Methods}
\label{sec:methods}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{fig/ps_torque.pdf}
    \caption{The \textit{upper panel} shows the evolution
    of the pattern speed. As expected, the bar in the \Nbody{} run slows down due
    to interactions between the bar and the dark matter halo. However, the bar in
    the SMUGGLE run does not slow down and instead remains at a constant pattern
    speed. The \textit{lower panel} shows the torque on the bar
    by different components. The solid lines show the torque exerted by the halo
    in both the \Nbody{} and SMUGGLE cases. The dashed line shows the torque
    exerted by the gas phase in the SMUGGLE run (there is no gas in the \Nbody{}
    run). Details on how these properties are calculated is given in the Methods
    section.}\label{fig:prop}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/Rb_A2.pdf}
    \caption{The evolution in bar length and strength. The \textit{upper panel}
    shows the evolution of the bar length. In the \Nbody{} case, the bar
    lengthens. This occurs because as the pattern speed drops, bar-like orbits
    at larger radii are possible. Stars are captured on these orbits,
    lengthening the bar. This process does not occur in the SMUGGLE cases since
    the bar pattern speed is not decreasing, and therefore the bar length
    remains constant. The bar strength, shown in the \textit{lower panel} is
    measured as the maximum of the second Fourier component divided by the
    zeroth Fourier component. We see that in the \Nbody{} case (blue) the bar
    strength increases with time, consistent with previous results showing that
    the bar strength increases as bars slow down. In the SMUGGLE case (orange),
    we see that the bar strength remains roughly constant, possibly slightly
    decreasing with time. This is also consistent with the expected relation
    between pattern speed and strength since the bar in this case is not slowing
    down.}
    \label{fig:strength}
\end{figure}
\subsection{Initial Conditions}
The initial setup of the galactic disk used in this work follows closely the
GALAKOS model \citep{2020ApJ...890..117D}, which uses a modified version of the
\texttt{MakeNewDisk} code \citep{2005MNRAS.361..776S}. The GALAKOS model has
three components - a radially exponential and vertically isothermal stellar
disk, and a stellar bulge and dark matter halo following a Hernquist profile
\citep{1990ApJ...356..359H}. All \Nbody{} runs in this work used the same setup
parameters as the GALAKOS disk, more details of which can be found in the
original paper.

@@

\section{Results}
\label{sec:results}

@@

@@

@@

@@

First, we note that \citet{2020MNRAS.491.3655G} found that the rotation
parameter \Rot{} positively correlates with gas fraction, such that galaxies
with higher gas fractions are rotating more slowly. However, it is not obvious
this is in tension with our result since the gas fraction of galaxies correlates
with other galactic properties \citep{2009ARAA..47..159B}. Furthermore, the
measurement of pattern speeds is a delicate process still prone to large errors.
\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/ps_late_start.pdf}
    \caption{Pattern speed evolution with a lower initial pattern speed. We
    tested the evolution of our system when gas is added to the \Nbody{} run at
    a later time, but with all other simulation parameters kept the same. The
    setup is therefore identical to our previous runs just with a bar that is
    larger, stronger, and with a lower pattern speed. We find that the pattern
    speed evolution is very similar to our fiducial case, except that the bar
    retains its original pattern speed. This indicates a mechanism which keeps
    the bar at its formation pattern speed, and that there is not a particular
    pattern speed which the system tends to.}
    \label{fig:snap700}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/ps_fgas.pdf}
    \caption{Pattern speed evolution with varying gas fractions. We explored the
    impact of lowering the initial gas surface density of our fiducial disk on
    the evolution of the pattern speed. The surface densities we tested of $20$,
    $15$, $10$, and $5\,\Msun/\textrm{kpc}^2$ correspond to initial gas
    fractions of $16\%$, $10\%$, $7\%$, and $4\%$. We find that initial surface
    densities $20$, $15$, and $10\,\Msun/\textrm{kpc}^2$ result in bars which
    remain at a constant pattern speed, while an initial surface density of
    $5\,\Msun/\textrm{kpc}^2$ results in a bar which slows down in roughly the
    same manner as the \Nbody{} case.}
    \label{fig:fgas}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/ps_GFM.pdf}
    \caption{Pattern speed evolution of a smooth ISM model. This evolution is
    shown for the fiducial disk in the \Nbody{} (blue), SMUGGLE (orange), and
    smooth ISM (red) cases. The smooth ISM model is an older model for the ISM
    which treats its multiphase nature in a subgrid fashion
    \citep{2003MNRAS.339..289S}. This fundamentally differs from the SMUGGLE
    model, which explicitly resolves the hot and cold phases of the ISM
    \citep{2019MNRAS.489.4233M}. The pattern speed in the smooth ISM case is
    broadly similar to the evolution in the SMUGGLE case. This shows that the
    stability of the pattern speed is not simply a result of our assumed model
    for the ISM.}
\label{fig:GFM}
\end{figure}
